When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in Contemporary America

### 2020-07-14 code/Submissionv1.Rmd
commit changes

## Military Credibility and Expertise

@@

##Political Polarization Today - Unique in America's History

@@

```{r include=TRUE, echo=FALSE}

@@

```{r fig2, fig.height=5, fig.width=7, fig.align="center", include=TRUE, echo=FALSE}
datatograph <- data.frame(
  year=c("1994","1996","1998","1999","2000","2003", "2004", "2006", "2007", "2010", "2013","2014", "2016", "2017", "2018", "2019", "2020"), count=c(1, 1,2, 3, 1, 1, 3, 3, 3, 1, 1, 1, 6, 5, 4, 6, 1)
)
ggplot(datatograph, aes(x =year, y =count)) + geom_bar(stat = "identity", width=.2) +
  labs(
    x = "Year",
    y = "Count",
    color = NULL,
    title = "Figure 2: # of Op-Eds Written By Military Retirees",
    subtitle = "On Topics of Social Policy, Domestic Policy, or Individual Figures"
  )
```

@@

This paper set out to examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years. Examining opinion pieces authored by retired military officers in major US newspapers revealed that these forces, together, have generated changes in how the military interacts with and what it says to the American public. The data examined demonstrates that the vast majority of op-eds published by retired military officers in the past fifty years exhibited professional standards. The data also showed, however, that retired military officers are arguing about topics that fall outside of the traditional umbrella of national security, and that they are expressing clear partisan preferences at a greater rate than they previously did. Moreover, both of these trends have increased under the past four years of Donald Trump's Presidency.

This paper also argued that it is the interaction of increasing institutional credibility and increasing political polarization that has the potential to harm the professional military ethic. Maintaining credibility and trust with the American public remains crucial for the military, but should a polarized political climate endure, which is suggested by the majority of polarization scholars, the military needs to also understand that credibility and trust can be abused if it is handled inappropriately. This is especially likely if military professionals, to include retirees, abuse the high level of credibility they currently enjoy to take up and advocate for partisan positions, or worse, to jockey with civilian leaders for power [@brooks_shaping_2008-1]. Unfortunately, as this paper illustrates, there are hints of this occurring. The letter signed by concerned graduates of West Point, which was clearly aimed at Secretaries Pompeo and Esper, and discussed at greater length earlier in this paper, is clear evidence of this.

Gaining a better understanding of the impacts of political polarization on the military profession remains a critical area for future research. Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public [@robinson_who_2019]. Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various "military outcomes," which may arguably include more important outcomes, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice. These outcomes and more, and the role of polarization in influencing them, deserve to be explored.

Finally, if one believes that today's polarization is different in degree and kind than previous eras of polarization, which has been argued in this paper, one must also recognize that the military will continue to face challenges as a result of political polarization. The events following the death of George Floyd certainly have revealed this.  
\singlespace
\pagebreak

>>>>

## _Military Credibility and Expertise_

@@

## _Political Polarization Today - Unique in America's History_

@@

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

@@

```{r fig2, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
datatograph <- data.frame(
  year=c("1994","1996","1998","1999","2000","2003", "2004", "2006", "2007", "2010", "2013","2014", "2016", "2017", "2018", "2019", "2020"), count=c(1, 1,2, 3, 1, 1, 3, 3, 3, 1, 1, 1, 6, 5, 4, 6, 1)
)
ggplot(datatograph, aes(x =year, y =count)) + geom_bar(stat = "identity", width=.2) +
  labs(
    x = "Year",
    y = "Count",
    color = NULL,
    title = "Figure 2: # of Op-Eds Written By Military Retirees",
    subtitle = "On Topics of Social Policy, Domestic Policy, or Individual Figures"
  )
```

@@

The goal of this paper was to explore and examine the ways in which political polarization and institutional credibility have impacted the professionalism of the military over the past fifty years. Examining the published opinion pieces authored by retired military officers in major US newspapers revealed that together, institutional credibility and political polarization have generated changes in the professionalism of some retired military officers, and more broadly, in how they as elites are communicating with the American public. Though the data examined demonstrates that the vast majority of op-eds published by retired military officers in the past fifty years exhibited professional standards, the data also showed two worrisome trends: that retired military officers are arguing about topics that fall outside of the traditional umbrella of national security, and that they are expressing clear partisan preferences at a greater rate than they previously did. Moreover, both of these trends have increased under the past four years of Donald Trump's Presidency.

This paper also developed a theory for how the interaction of increasing institutional credibility and increasing political polarization has potential to harm the professional military ethic. Maintaining credibility and trust with the American public remains crucial for the military, but should a polarized political climate endure, which is suggested by the majority of American politics scholars who study polarization, the military needs to also understand that credibility and trust can be abused if it is handled inappropriately - a legitimate risk in a polarized environment. This is especially likely if military professionals, to include retirees, abuse the high level of credibility they currently enjoy to take up and advocate for partisan positions, or worse, to jockey with civilian leaders for power [@brooks_shaping_2008-1]. Unfortunately, as this paper illustrates, there are hints of this occurring. The letter signed by concerned graduates of West Point, which was clearly aimed at Secretaries Pompeo and Esper, and discussed at greater length earlier in this paper, is clear evidence of this.

Gaining a better understanding of the impacts of political polarization on the military profession remains a critical area for future research. Other scholars such as Michael Robinson have explored the specific ways that polarization is impacting the interactions between the military and the public [@robinson_who_2019]. Though this paper focused on the ways that polarization may impact professional military behavior, it recognized but stopped short of discussing polarization's subsequent impact on various "military outcomes," which may arguably include more decisive events, such as achieving victory or defeat in battle, or the ability to formulate sound strategic advice. These outcomes and more, and the role of polarization in influencing them, deserve to be explored further.

Finally, if one believes that today's polarized climate in America is different in degree and kind than previous eras of polarization, which has been a supporting claim of this paper, one must recognize that the military will face continued institutional challenges stemming from polarization's underlying causes, and so will individual military leaders. The most portending aspect regarding today's polarizing climate is that fundamental values are at stake, values that speak directly to the heritage of the United States as one nation under God and to the military as the nation's guarantor of individual rights and freedoms. As a non-partisan institution that should theoretically be free from partisan entanglement, the actions of the military will remain under close scrutiny by the American public. In this type of environment, in some sense, the military is doomed: regardless of how the institution or its leaders act or respond to pressing polarized debates, some will accuse it of failing to do its duty and to exercise genuine loyalty. It is during these tough moments that another soldierly virtue - moral courage - will perhaps stand out as the chief attribute of military leaders, now and in the years to come. 
\singlespace
\pagebreak
